### 2021-06-18 paper/report.tex
Went through draft once more.

Since early 2020,\comment[id=J]{Put stronger focus on time dependent rapid test
sensitivity (30\% before day of onset of infectiousness)} the CoViD-19 pandemic has
presented an enormous challenge to humanity on many dimensions. The development of
highly effective vaccines holds the promise of containment in the medium term. However,
most countries find themselves many months---and often years---away from reaching
vaccination-induced herd immunity \citep{Swaminathan2021}.\comment[id=HM]{Need to find
some more scientific citation before submission} In the meantime, it is of utmost
importance to employ an effective mix of strategies for containing the virus. The most
frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
contacts between individuals. While this has allowed some countries to sustain
equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
theoretical equilibrium at low case numbers which is sustained with
test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
over time. Containment measures have become increasingly diverse and now include
testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
influence of seasonal patterns or more infectious virus strains are well understood in
quantitative terms.

@@

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:model_contacts_infections}).\footnote{A detailed comparison with other
approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
model most closely related to ours is described in \citet{Hinch2020}.} Each contact
between an infectious individual and somebody susceptible to the disease bears the risk
of transmitting the virus. Contacts occur in up to four networks: Within the household,
at work, at school, or in other settings (leisure activities, grocery shopping, medical
appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
applications can take the population and household structure from census data and the
network-specific frequencies of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, contacts are
differentiated by age and geographical location see Section~\ref{sub:number_of_contacts}
of the Supplementary Material.

@@

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death as for example also modeled in \cite{Grimm2021}.
The probabilities of transitioning between these states depend on age; their duration is
random within intervals calibrated to medical literature (for a detailed description see
Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
independent of age \citep{Jones2021}.

@@

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and estimate the model's many
parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
materials for an overview.} Contact diaries yield pre-pandemic distributions of contacts
for different contact types and their assortativity by age group. Mobility data is used
to model the evolution of work contacts. School and daycare policies can be incorporated
directly from official directives. Administrative records on the number of tests,
vaccinations by age and region, and the prevalence of virus strains are generally
available. Surveys may ask about test offers, propensities to take them up, and past
tests. Other studies' estimates of the seasonality of infections can be incorporated
directly. The remaining parameters---most notably, these include infection probabilities
by contact network and the effects of NPIs, see Supplementary
Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
matches features of the data \citep[see][for the general method, also described in
Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make reference
more specific once Appendix has new structure}. We keep the number of free parameters
low in order to avoid overfitting. In our application, these features include  official
case numbers for each age group and region, deaths, and the share of the B.1.1.7 strain.

@@

By this time, the set of policy instruments had become much more diverse. Around the
turn of the year, the first people were vaccinated with a focus on older age groups and
medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}). By the end of May, just
over 40\% had received at least one dose of a vaccine. Around the same time, rapid tests
started to replace regular PCR tests for staff in many medical and nursing facilities.
These had to be administered by medical doctors or in pharmacies. At-home tests approved
by authorities became available in mid-March, rapid test centers were opened and one
test per person and week was made available free of charge. Depending on the state,
customers were only allowed to enter certain stores with a recent negative rapid test
result. These developments are characteristic of many countries: The initial focus on
NPIs to slow the spread of the disease has been accompanied by vaccines and a growing
acceptance and use of rapid tests. At broadly similar points in time, novel strains of
the virus have started to pose additional challenges.

We draw simulated samples of agents from the population structure in September 2020 and
use the model to predict recorded infection rates until the end of May 2021. See
Supplementary Material~\ref{sub:initial_conditions} for a detailed description of this
procedure. The blue line in Figure~\ref{fig:aggregated_fit} shows our model's
predictions are very close to officially recorded cases in the aggregate. This is also
true for infections by age and geographical region, which are shown in the supplementary
materials (Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively).

@@

    \floatfoot{\noindent \textit{Note:} The blue line in
        Figure~\ref{fig:2021_scenarios_recorded} is the same as in
        Figure~\ref{fig:aggregated_fit} and refers to our baseline scenario, so does the
        blue line in Figure~\ref{fig:2021_scenarios_newly_infected}. The red lines refer
        to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
        variant is introduced in the same way; vaccinations, rapid tests, and
        seasonality remain at their January levels. The other scenarios turn these three
        factors on one-by-one. The decompositions in
        Figures~\ref{fig:2021_scenarios_decomposition} and
        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values which
        are explained more thoroughly in Appendix~\ref{sub:shapley_value}}
\end{figure}

@@

The largest effect---almost one half when considering the decompositions---comes from
rapid testing. Here, it is crucial to differentiate between recorded cases and actual
cases. Additional testing means that infections are recorded, which would otherwise
remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is
large. Until late April, recorded cases are higher in the scenario with rapid testing
alone when compared to the setting where none of the three mechanisms are turned on.
Compared to the scenario with vaccinations only, this point is reached around mid-May
and it would be June for the comparison with the seasonality-only scenario. The effect
on total cases, however, is visible immediately in
Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only a small
fraction of the population performed weekly rapid tests in March,\footnote{4\% to 6\% of
respondents to the COSMO study report having done a rapid test during the first week of
March; 8\% in the model have ever done a rapid test at that point.}\comment[id=HM]{would
be great to have total tests performed in March, model vs. data is not important here.
Just first week of March looks too small} new infections on April 1 would be reduced by
53\% relative to the scenario without vaccinations, rapid tests, or seasonality.

>>>>

Since early 2020,\comment[id=J]{Put stronger focus on time dependent rapid test
sensitivity (30\% before day of onset of infectiousness)} the CoViD-19 pandemic has
presented an enormous challenge to humanity on many dimensions. The development of
highly effective vaccines holds the promise of containment in the medium term. However,
most countries find themselves many months---and often years---away from reaching
vaccination-induced herd immunity \citep{Swaminathan2021}.\comment[id=HM]{Need to find
some more scientific citation before submission} In the meantime, it is of utmost
importance to employ an effective mix of strategies for containing the virus. The most
frequent initial response was a set of non-pharmaceutical interventions (NPIs) to reduce
contacts between individuals. While this has allowed some countries to sustain
equilibria with very low infection numbers,\footnote{See \citet{Contreras2021} for a
theoretical equilibrium at low case numbers which is sustained with
test-trace-and-isolate policies.} most have seen large fluctuations of infection rates
over time. Containment measures have become increasingly diverse and now include rapid
testing, more nuanced NPIs, and contact tracing. Neither these policies' effect nor the
influence of seasonal patterns or of more infectious virus strains are well understood
in quantitative terms.

@@

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:model_contacts_infections}).\footnote{A detailed comparison with other
approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
model most closely related to ours is described in \citet{Hinch2020}.} Each contact
between an infectious individual and somebody susceptible to the disease bears the risk
of transmitting the virus. Contacts occur in up to four networks: Within the household,
at work, at school, or in other settings (leisure activities, grocery shopping, medical
appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
applications can take the population and household structure from census data and the
network-specific frequencies of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, contacts are
differentiated by age and geographical location (see Section~\ref{sub:number_of_contacts}
of the Supplementary Material).

@@

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death \citep[similar
to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The probabilities of
transitioning between these states depend on age; their duration is random within
intervals calibrated to medical literature (for a detailed description see
Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
independent of age \citep{Jones2021}.

@@

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and estimate the model's many
parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
materials for an overview.} Contact diaries yield pre-pandemic distributions of contacts
for different contact types and their assortativity by age group. Mobility data is used
to model the evolution of work contacts. School and daycare policies can be incorporated
directly from official directives. Administrative records on the number of tests,
vaccinations by age and region, and the prevalence of virus strains are generally
available. Surveys may ask about test offers, propensities to take them up, and past
tests. Other studies' estimates of the seasonality of infections can be incorporated
directly. The remaining parameters---most notably, these include infection probabilities
by contact network and the effects of NPIs, see Supplementary
Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
matches features of the data \citep[see][for the general method, also described in
Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make reference
more specific once Appendix has new structure} In our application, we keep the number of
free parameters low in order to avoid overfitting. The data features to be matched
include official case numbers for each age group and region, deaths, and the share of
the B.1.1.7 strain.

@@

By this time, the set of policy instruments had become much more diverse. Around the
turn of the year, the first people were vaccinated with a focus on older age groups and
medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}). By the end of May, just
over 40\% had received at least one dose of a vaccine. Around the same time, rapid tests
started to replace regular PCR tests for staff in many medical and nursing facilities.
These had to be administered by medical doctors or in pharmacies. At-home tests approved
by authorities became available in mid-March, rapid test centers were opened, and one
test per person and week was made available free of charge. In several states, customers
were only allowed to enter certain stores with a recent negative rapid test result.
These developments are characteristic of many countries: The initial focus on NPIs to
slow the spread of the disease has been accompanied by vaccines and a growing acceptance
and use of rapid tests. At broadly similar points in time, novel strains of the virus
have started to pose additional challenges.

We draw simulated samples of agents from the population structure in September 2020 and
use the model to predict recorded infection rates until the end of May 2021. See
Supplementary Material~\ref{sub:initial_conditions} for a detailed description of this
procedure. The blue line in Figure~\ref{fig:aggregated_fit} shows that our model's
predictions are very close to officially recorded cases in the aggregate. This is also
true for infections by age and geographical region, which are shown in the supplementary
materials (Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively).

@@

    \floatfoot{\noindent \textit{Note:} The blue line in
        Figure~\ref{fig:2021_scenarios_recorded} is the same as in
        Figure~\ref{fig:aggregated_fit} and refers to our baseline scenario, so does the
        blue line in Figure~\ref{fig:2021_scenarios_newly_infected}. The red lines refer
        to a situation where NPIs evolve as in the baseline scenario and the B.1.1.7
        variant is introduced in the same way; vaccinations, rapid tests, and
        seasonality remain at their January levels. The other scenarios turn these three
        factors on one-by-one. The decompositions in
        Figures~\ref{fig:2021_scenarios_decomposition} and
        \ref{fig:2021_scenarios_decomposition_tests} are based on Shapley values, which
        are explained more thoroughly in Appendix~\ref{sub:shapley_value}}
\end{figure}

@@

The largest effect---almost one half when considering the decompositions---comes from
rapid testing. Here, it is crucial to differentiate between recorded cases and actual
cases. Additional testing means that additional infections will be recorded which would
otherwise remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this
effect is large. Until late April, recorded cases are higher in the scenario with rapid
testing alone when compared to the setting where none of the three mechanisms are turned
on. Compared to the scenario with vaccinations only, this point is reached around
mid-May and it would be June for the comparison with the seasonality-only scenario. The
effect on total cases, however, is visible immediately in
Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\%  of the
population performed weekly rapid tests in March on average, new infections on April 1
would be reduced by 53\% relative to the scenario without vaccinations, rapid tests, or
seasonality.
